|

### 2021-03-07 manuscript_science.Rmd
edits: manuscript, introduction

**Abstract:** Larger ecosystems support more species; this ubiquitous ecological pattern is the foundation of current conservation schemes. However, many ecosystems possess a complex spatial structure that cannot be represented by area, and the role of such complexity in scaling biodiversity is largely unknown. Here, we use theory and extensive fish community data from two distinct geographic regions (Japan and United States) to show that ecosystem size and complexity dictate riverine biodiversity. We found that larger and more branched 'complex' river networks harbored greater species richness due to increased space and environmental heterogeneity. The complexity effect was comparable to the size effect, and this pattern emerged regardless of ecological contexts. The dual control of biodiversity may be a pervasive feature that has far-reaching implications for biodiversity conservation.

@@

Many ecosystems, however, possess a complex spatial structure that cannot be represented by area - a dimension referred to as scale-invariant complexity [@turner2015; @rodríguez-iturbe2001]. Such complexity is evident in branching ecosystems, including rivers, trees, and mountain ranges, to name just a few [@rodríguez-iturbe2001]. Geomorphic or biological processes generate a pronounced self-similarity in complex branching patterns such that the part and the whole look alike [@rodríguez-iturbe2001]. Even though the branching structure is independent of spatial scale, it forms a physical template that dictates habitat diversity and dispersal corridors for living organisms [@grant2007]. Limited, but accumulating evidence suggests that classical metapopulation and metacommunity theories cannot predict ecological dynamics driven by branching structure [@fagan2002; @grant2010; @brown2010; @henriquessilvarenato], and this recognition has led to recent developments of spatial theories devoted to complex branching ecosystems [@rinaldo2018]. For example, these studies have highlighted key roles of branching structure in driving local biodiversity patterns, such as increased species richness at merging points of branches [@carrara2012]. However, most empirical research has explored the consequences of branching complexity for local community structure [@ers2019; @carrara2012] or has relied solely on theoretical arguments with limited replications of branching architecture [@holt2018]. At present, we lack a synthetic framework integrating the influences of ecosystem size and complexity on biodiversity at the metacommunity level. Filling this knowledge gap may provide common ground for achieving successful conservation in spatially complex ecosystems, where accelerated species loss threatens the delivery of ecosystem services [@dudgeon2006].

Here, we hypothesize that ecosystem size and complexity dictate biodiversity patterns in rivers - a prime example of complex branching ecosystems. Individual streams and rivers flow through different landscapes with distinct geological and climatic backgrounds, serving as a spatial unit of unique in-stream environments [@brown2010; @peterson2013; @moore2015; @terui2018]. The recurrent merging of diverse tributaries ultimately forms a fluvial network with fractal branching patterns [@rodríguez-iturbe2001]. As such, the complexity of branching structure, which we define here as the probability of branching per unit river distance [@terui2018; @yeakel2014], may represent a macro-scale control of the ecosystem's habitat heterogeneity (habitat diversity per unit area) [@grant2007; @moore2015; @terui2018]. Meanwhile, ecosystem size (watershed area) should determine the metacommunity size and total habitat diversity (area x heterogeneity), two factors that regulate biodiversity at the metacommunity level [@hubbell2001; @iwasaki2012]. We predict that ecosystem size and branching complexity increase watershed-scale species richness ($\gamma$ diversity) by enhancing local species richness ($\alpha$ diversity) and/or spatial difference of species composition ($\beta$ diversity) under different ecological scenarios. The present study combines theory and empirical analysis of extensive community data from two different regions of the globe to provide crucial insights into how ecological communities are structured in complex branching networks.

>>>>

**Abstract:** Larger ecosystems support more species; this ubiquitous phenomenon is the foundation of current conservation schemes. However, many ecosystems possess a complex spatial structure that cannot be represented by area, and the role of such complexity in scaling biodiversity is largely unknown. Here, we use theory and extensive fish community data from two distinct geographic regions (Japan and United States) to show that ecosystem size and complexity dictate riverine biodiversity. We found that larger and more branched 'complex' river networks harbored greater species richness due to increased space and environmental heterogeneity. The complexity effect was comparable to the size effect, and this pattern has emerged regardless of ecological contexts. The dual control of biodiversity may be a pervasive feature that has far-reaching implications for biodiversity conservation.

@@

Many ecosystems, however, possess a complex spatial structure that cannot be represented by area - a dimension referred to as scale-invariant complexity [@turner2015; @rodríguez-iturbe2001]. Such complexity is evident in branching ecosystems, including rivers, trees, and mountain ranges, to name just a few [@rodríguez-iturbe2001]. Geomorphic or biological processes generate a pronounced self-similarity in complex branching patterns such that the part and the whole look alike [@rodríguez-iturbe2001]. Even though the branching structure is independent of spatial scale, it forms a physical template that dictates habitat diversity and dispersal corridors for living organisms [@grant2007]. Limited, but accumulating evidence suggests that classical metapopulation and metacommunity theories cannot predict ecological dynamics driven by branching structure [@fagan2002; @grant2010; @brown2010; @henriquessilvarenato], and this recognition has led to recent developments of spatial theories devoted to complex branching ecosystems [@rinaldo2018]. For example, these studies have highlighted key roles of branching structure in driving local biodiversity patterns, such as increased species richness at merging points of branches [@carrara2012]. However, most empirical research has explored the consequences of branching complexity for local community structure [@ers2019; @carrara2012] or has relied solely on theoretical arguments with limited replications of branching architecture [@holt2018]. The lack of a metacommunity-level framework integrating the influences of ecosystem size and complexity hinders an objective assessment of their roles in controlling biodiversity patterns. Filling this knowledge gap may provide common ground for achieving successful conservation in spatially complex ecosystems, where accelerated species loss threatens the delivery of ecosystem services [@dudgeon2006].

Here, we show that ecosystem size and complexity play comparable roles in controlling biodiversity patterns in rivers - a prime example of complex branching ecosystems. Individual streams and rivers flow through different landscapes with distinct geological and climatic backgrounds, serving as a spatial unit of unique in-stream environments [@brown2010; @peterson2013; @moore2015; @terui2018]. The recurrent merging of diverse tributaries ultimately forms a fluvial network with fractal branching patterns [@rodríguez-iturbe2001]. As such, the complexity of branching structure, which we define here as the probability of branching per unit river distance [@terui2018; @yeakel2014], may represent a macro-scale control of the ecosystem's habitat heterogeneity (habitat diversity per unit area) [@grant2007; @moore2015; @terui2018]. Meanwhile, ecosystem size (watershed area) should determine the metacommunity size and total habitat diversity (area x heterogeneity), two factors that regulate biodiversity at the metacommunity level [@hubbell2001; @iwasaki2012]. We predict that ecosystem size and branching complexity increase watershed-scale species richness ($\gamma$ diversity) by enhancing local species richness ($\alpha$ diversity) and/or spatial difference of species composition ($\beta$ diversity) under different ecological scenarios. The present study combines theory and empirical analysis of extensive community data from two different regions of the globe to provide crucial insights into how ecological communities are structured in complex branching networks.

### 2021-03-07 supplementary_information_science.Rmd
edits: manuscript, introduction

  # for bayes factor
  source(here::here('empirical/analysis_lm_bf.R'))

  # Hokkaido data
  dat_hkd_site <- read_csv(here::here('empirical/data_out/data_hkd_site.csv'))

  # Midwest data
  dat_mw_site <- read_csv(here::here('empirical/data_out/data_mw_site.csv')) %>% 
    mutate(Year = format(as.Date(Date, format = '%m/%d/%Y'), "%Y"))

  # sensitivity tables
  source(here::here('theory/table_sensitivity_analysis.R'))

@@

## Empirical data analysis

@@

We estimated $\alpha$, $\beta$, and $\gamma$ diversity at each watershed. $\gamma$ diversity is the total species richness at each watershed. Since it is impossible to sample all species present in each watershed, we used asymptotic species richness as a proxy for $\gamma$ diversity. Asymptotic species richness can be interpreted as the estimated true species richness in a given area when estimated using spatial replicates of local community data within a metacommunity [@chao2014]. We estimated asymptotic species richness using sampling sites located in the same watershed (i.e., spatial replicates within a metacommunity) via the R package 'iNEXT' [@hsieh2016]. We assessed the completeness of the observed species richness based on sample coverage, which is the proportion abundances of observed species to the whole community [@chao2014]. We excluded watersheds whose sample coverage was below 0.9 because the estimates of asymptotic species richness may contain non-negligible statistical uncertainty. $\alpha$ diversity was estimated by taking an average of local species richness at each site. $\beta$ diversity was estimated by dividing the estimated $\gamma$ diversity by $\alpha$ diversity [@whittaker1960].

@@

We developed linear regression models to examine the effects of macro-scale drivers on diversity metrics. Log-transformed diversity metrics ($\alpha$, $\beta$ and $\gamma$ diversity) at watershed $w$, $log_{10}~y_w$, were assumed to follow a normal distribution as $log_{10}~y_w \sim Normal(\mu_w, \sigma^2)$. The parameter $\mu_w$ was related to linear predictors, including log-transformed watershed area (ecosystem size; $log_{10}~WA$), log-transformed branching probability (ecosystem complexity; $log_{10}~P_b$), region (coded as Hokkaido = 0, Midwest = 1; $region$), air temperature, precipitation, the fraction of agricultural land use (logit-transformed), and dam density. Although our primary focus of this analysis is to assess the effects of watershed area and branching probability, we included climate and land use variables to control their effects statistically. We did not include the fractions of urban and forest land use in the statistical models because these variables were strongly correlated with the fraction of agricultural land use (**Figure S19**). Similarly, we did not use mean elevation in the statistical models because it showed a moderate correlation with air temperature (**Figure S19**). Correlations for other combinations of explanatory variables were weak to moderate.

@@

>>>>

  # Hokkaido data
  dat_hkd <- read_csv(here::here("empirical/data_out/data_hkd.csv")) %>% mutate(region = "hokkaido")
  dat_hkd_site <- read_csv(here::here('empirical/data_out/data_hkd_site.csv'))

  # Midwest data
  dat_mw <- read_csv(here::here("empirical/data_out/data_mw.csv")) %>% mutate(region = "midwest")
  dat_mw_site <- read_csv(here::here('empirical/data_out/data_mw_site.csv')) %>% 
    mutate(Year = format(as.Date(Date, format = '%m/%d/%Y'), "%Y"))

  # tables

  ## sensitivity
  source(here::here('theory/table_sensitivity_analysis.R'))

@@

@@

We estimated $\alpha$, $\beta$, and $\gamma$ diversity at each watershed. $\gamma$ diversity is the total species richness at each watershed. Since it is impossible to sample all species present in each watershed, we used asymptotic species richness as a proxy for $\gamma$ diversity. Asymptotic species richness can be interpreted as the estimated true species richness in a given area when estimated using spatial replicates of local community data within a metacommunity [@chao2014]. We estimated asymptotic species richness using sampling sites located in the same watershed (i.e., spatial replicates within a metacommunity) via the R package 'iNEXT' [@hsieh2016]. We assessed the completeness of the observed species richness based on sample coverage, which is the total proportional abundances of observed species to the whole metacommunity [@chao2014]. We excluded watersheds whose sample coverage was below 0.9 because the estimates of asymptotic species richness may contain non-negligible statistical uncertainty. $\alpha$ diversity was estimated by taking an average of local species richness at each site. $\beta$ diversity was estimated by dividing the estimated $\gamma$ diversity by $\alpha$ diversity [@whittaker1960].

@@

We developed robust regression models with an M-estimation method to examine the effects of macro-scale drivers on diversity metrics. We employed robust regression models given the uncertainty of diversity estimates (see **Fish community data**). Log-transformed diversity metrics ($\alpha$, $\beta$ and $\gamma$ diversity) at watershed $w$, $log_{10}~y_w$, were assumed to follow a normal distribution as $log_{10}~y_w \sim Normal(\mu_w, \sigma^2)$. The parameter $\mu_w$ was related to linear predictors, including log-transformed watershed area (ecosystem size; $log_{10}~WA$), log-transformed branching probability (ecosystem complexity; $log_{10}~P_b$), region (coded as Hokkaido = 0, Midwest = 1; $region$), air temperature, precipitation, the fraction of agricultural land use (logit-transformed), and dam density. Although our primary focus of this analysis is to assess the effects of watershed area and branching probability, we included climate and land use variables to control their effects statistically. We did not include the fractions of urban and forest land use in the statistical models because these variables were strongly correlated with the fraction of agricultural land use (**Figure S19**). Similarly, we did not use mean elevation in the statistical models because it showed a moderate correlation with air temperature (**Figure S19**). Correlations for other combinations of explanatory variables were weak to moderate.

@@

### Data selection criteria
